---
title: "RExercise#2_Virgo"
Author: "Erlle Benjamin R. Virgo"
output: html_document
date: "2026-09-30"
---


# PART A - ARITHMETIC OPERATIONS 

##### TODO: Store 17 in a variable called a, and 5 in a variable called b
##### Then print the results of: a + b, a - b, a * b, a / b

```{r}
a <- 17
b <- 5

a + b
a - b
a * b
a / b
```
#### TODO: Using the same a and b, compute and print a %% b and a %/% b
#### a %% b = remainder of a / b
#### a %/% b = whole-number (integer) division
  
```{r}
a %% b 
a %/% b
```
##### TODO: Print 2 raised to the power of 10 using the ^ operator
```{r}
2^10
```

##### Question #1: Run 17 / 5, 17 %% 5, and 17 %/% 5 side by side. In your own words, explain how these three results relate to each other mathematically.

```{r}
17 / 5
17 %% 5
17 %/% 5
# all three results comes from the same division
```
##### Question #2: Predict the output of 10 %% 3 and 10 %/% 3 before running the code. Were you correct? If not, explain what you misunderstood.

```{r}
10 %% 3
10 %/% 3
```



#### PART B - BASIC DATA TYPES: typeof(), class(), and mode()
```{r}
x <- 42
y <- 42L        
w <- TRUE
v <- 3 + 2i    
      
typeof(x)
typeof(y)
typeof(w)
typeof(v)

class(x)
class(y)
class(w)
class(v)

mode (x)
mode (y)
mode (w)
mode (v)

```
###### Question #3): x <- 42 and y <- 42L look almost identical, but typeof(x) and typeof(y) give different answers. Explain the difference in your own words.

```{r}
# x <- 42 makes a double, while y <- 42L makes an integer. So they print the same value which is 42, but R stores them differently
```
###### Question #4: Is typeof() ever different from class() and from mode() for the same object? Look at your results above and explain any case where this happens (or, if you did not see one, explain what you would expect for a matrix or factor).
```{r}
# Yes they can be different. For x <- 42, typeof(x) it says "double", but class(x) and mode(x) both say "numeric" .
```

#### PART C - Type conversion (20 min)

```{r}
num_string <- "123"
num <- as.numeric(num_string)
typeof(num)
class(num)
mode(num)
```
```{r}
as.logical(0)
as.logical(1)
```

```{r}
bad_string <- "hello123"
as.numeric(bad_string)
```


#### Part D - Strings, paste(), and paste0()




